Appendix 1: Preregistered analysis with simulated data

### 2025-04-23 preregistration_code.Rmd
missing fixes

@@

```{r}
load_data()
if (exists("data_simulated")) {
  data <- data_simulated
  rm(data_simulated)
}
# sometimes I get strange errors about C stack which can be solved with this
#data <- as.data.frame(data)
fitmodels <- FALSE
if (!fitmodels) {
  print("Loading previously fitted models")
  load(file=here("prereg_models.rdata"))
  fit.po02ss <- models$fit.po02ss
  fit.po03ss <- models$fit.po03ss
  fit.po04ss <- models$fit.po04ss
  fit.po02 <- models$fit.po02
  fit.po03 <- models$fit.po03
  fit.po04 <- models$fit.po04
}

@@

Check is missingness related to data accumulation.

@@

Check is missingness related to major endpoint of the study i.e. intention to donate blood.

@@

if (fitmodels) {
  fit.po04ss <- sem(
    po02ss.model,  
    estimator = "ML", # 
    data = data.fit2, # data that is completely numeric
    auto.fix.first = FALSE,
    std.lv = TRUE,
    missing = "fiml",
    fixed.x = F 
  )
  models[['fit.po04ss']] <- fit.po04ss
}

@@

if (fitmodels) {
  fit.po04 <- sem(
    po02.model,  
    estimator = "ML", # 
    data = data.fit2, # data that is completely numeric
    auto.fix.first = FALSE,
    std.lv = TRUE,
    missing = "fiml",
    fixed.x = F 
  )
  models[['fit.po04']] <- fit.po04
}

@@

```{r}
lavaanPlot(
  model = fit.po04, 
   labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))

@@

```{r}
data.forest  <- fit.po03.best %>% 
  mutate(estmet = "Factor analysis\nbootstrapped") %>%  
  #rename(est=est.std) %>% 
  filter(op == "~" | op == ":=") %>% 
  bind_rows(
    fit.po02.est %>% mutate(estmet = "Factor analysis") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":="),
    fit.po04.est %>% mutate(estmet = "Factor analysis\nimputed") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") 
  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs)
    ) %>% 
  arrange(estmet, term)

data.forest  <- fit.po03ss.best %>% 
  rename(est=est.std) %>% 
  mutate(estmet = "Sum score\nbootstrapped") %>%  
  filter(op == "~"| op == ":=") %>% 
  bind_rows(
    fit.po02ss.est %>% mutate(estmet = "Sum score") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") ,
    fit.po04ss.est %>% mutate(estmet = "Sum score\nimputed") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=")

  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs)
    ) %>% 
  arrange(estmet, term) %>% 
  bind_rows(data.forest) %>% 
  mutate(
    lhs =str_replace_all(lhs,"ss_",""),
    rhs =str_replace_all(rhs,"ss_",""),
    term =str_replace_all(term,"ss_",""),
    term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
    term =str_replace_all(term,"bd02","blooddonor:"),
    term =str_replace_all(term,"blooddonor:","blooddonor"),
    term =str_replace_all(term,"^:",""),
    estmet = as.factor(estmet),
    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(estmet)),
                             levels=levels(estmet))
  )

pos <- position_nudge(y=as.numeric(as.factor(data.forest$estmet))/5 -0.5)

data.forest %>% 
  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=estmet))+
  geom_vline(xintercept=0,color="grey") +
  geom_linerange(position=pos)+
  geom_point(shape=21,size=3,position=pos)+
  scale_fill_discrete(na.value=NA, guide ="none") + 
  labs(x="Standardised effect size",y="Relationship") +
  labs(color="Model") +
  geom_stripped_rows(color=NA) + theme_classic()

@@

@@

round(
  sapply(list(SS=fit.po02ss,SSB=fit.po03ss,FA=fit.po02,FAB=fit.po03 ),
         function(x) fitmeasures(x,c("chisq","df","cfi","rmsea","srmr"))
         )
,3)

@@

srmr Standardized Root Mean Square Residual, "Should be < .08", smaller is better

>>>>

# renv::status() to check 
# renv::snapshot() to add used and install to renv

@@

```{r}
load_data()
if (exists("data_simulated")) {
  data <- data_simulated
  rm(data_simulated)
}
# sometimes I get strange errors about C stack which can be solved with this
#data <- as.data.frame(data)
fitmodels <- FALSE # TRUE for fitting models
if (!fitmodels) {
  print("Loading previously fitted models")
  load(file=here("prereg_models.rdata"))
  fit.po02ss <- models$fit.po02ss
  fit.po03ss <- models$fit.po03ss
  fit.po04ss <- models$fit.po04ss
  fit.po02 <- models$fit.po02
  fit.po03 <- models$fit.po03
  fit.po04 <- models$fit.po04
}

@@

Check if missingness is related to data accumulation.

@@

Check if missingness is related to major endpoint of the study i.e. intention to donate blood.

@@

if (fitmodels) {
  fit.po04ss <- sem(
    po02ss.model,  
    estimator = "ML", # 
    data = data.fit2, # data that is completely numeric
    auto.fix.first = FALSE,
    std.lv = TRUE,
    missing = "fiml",
    fixed.x = FALSE 
  )
  models[['fit.po04ss']] <- fit.po04ss
}

@@

if (fitmodels) {
  fit.po04 <- sem(
    po02.model,  
    estimator = "ML", # 
    data = data.fit2, # data that is completely numeric
    auto.fix.first = FALSE,
    std.lv = TRUE,
    missing = "fiml",
    fixed.x = FALSE 
  )
  models[['fit.po04']] <- fit.po04
}

@@

```{r}
lavaanPlot(
  model = fit.po04, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))

@@

### Effect sizes

data.forest  <- fit.po03.best %>% 
  mutate(estmet = "Bootstrapped"
         ) %>%  
  #rename(est=est.std) %>% 
  filter(op == "~" | op == ":=") %>% 
  bind_rows(
    fit.po02.est %>% mutate(estmet = "Robust") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":="),
    fit.po04.est %>% mutate(estmet = "Imputed") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") 
  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs),
    linetype = "Latent factor"
    ) %>% 
  arrange(estmet, term)

data.forest  <- fit.po03ss.best %>% 
  rename(est=est.std) %>% 
  mutate(estmet = "Bootstrapped") %>%  
  filter(op == "~"| op == ":=") %>% 
  bind_rows(
    fit.po02ss.est %>% mutate(estmet = "Robust") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") ,
    fit.po04ss.est %>% mutate(estmet = "Imputed") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=")

  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs),
    linetype = "Sum score"
    ) %>% 
  arrange(estmet, term) %>% 
  bind_rows(data.forest) %>% 
  mutate(
    lhs =str_replace_all(lhs,"ss_",""),
    rhs =str_replace_all(rhs,"ss_",""),
    term =str_replace_all(term,"ss_",""),
    term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
    term =str_replace_all(term,"bd02","blooddonor:"),
    term =str_replace_all(term,"blooddonor:","blooddonor"),
    term =str_replace_all(term,"^:",""),
    estmet = as.factor(estmet),
    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(estmet)),
                             levels=levels(estmet))
  ) %>% 
  mutate(
    mtype = str_c(estmet,"-",linetype)
  )

pos <- position_nudge(y=as.numeric(as.factor(data.forest$mtype))/7 -0.5)

pf <- data.forest %>% 
  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=estmet,linetype = linetype))+
  geom_vline(xintercept=0,color="grey") +
  geom_linerange(position=pos)+
  geom_point(shape=21,size=3,position=pos,alpha=0.7)+
  scale_fill_discrete(na.value=NA, guide ="none") + 
  labs(x="Standardised effect size",y="Relationship") +
  labs(color="Model type", linetype="Variable type") +
  geom_stripped_rows(color=NA) + theme_classic()

pf
```

```{r}
ggsave(pf,
       filename = here("figures/forest01.png"),
       height = 140,
       width = 180,
       units = "mm"
         )

ggsave(pf,
       filename = here("figures/forest01.pdf"),
       height = 140,
       width = 180,
       units = "mm"
         )

@@

## Indices of goodness of fit

@@

gfi<- round(
  sapply(
    list(
      SS=fit.po02ss,
      SSB=fit.po03ss,
      SSI = fit.po04ss,
      FA=fit.po02,
      FAB=fit.po03, 
      FAI=fit.po04
      ),
         function(x) fitmeasures(x,c("chisq","df","pvalue","cfi","rmsea","srmr"))
         )
,3)

gfi
```

@@

srmr Standardized Root Mean Square Residual, "should be < .08", smaller is better

NB. Sum score and latent factor models use different data and are not nested hence they cannot be compared with these metrics.
